### 2020-09-23 doc/ACCAHA_BP_groups.Rmd
cleaning manuscript for consistency

---
title: "Predicted cardiovascular risk and blood pressure for Americans with diabetes, chronic kidney disease, and \u226565 years of age"
bibliography: refs.bib
csl: jacc.csl
output: 
  officedown::rdocx_document:
    reference_docx: style_manuscript.docx
---

```{r setup, include=FALSE}

# misc functions for the paper

# refers to tables and figures
refer_to <- function(label){
  report$pre_cap[report$reference == label]
}

# fixes list names with symbols
clean_list_names <- function(named_list, case){

  new_names <- make_clean_names(string = names(named_list),
                                case = case)

  names(named_list) <- new_names

  named_list

}

# adds counter to abbreviation key
add_counter <- function(data, index){
  data$counter[index] <- data$counter[index] + 1
  data
}

knitr::opts_chunk$set(echo = FALSE)

loadd(exams)
loadd(current_analysis)
loadd(inline)
loadd(report)
loadd(fasted_hrs_lower)
loadd(fasted_hrs_upper)
loadd(gluc_cutpoint_fasted)
loadd(gluc_cutpoint_fed)
loadd(hba1c_cutpoint)
loadd(egfr_cutpoint)
loadd(acr_cutpoint)

nhanes_years <- exams %>%
  map_chr(~glue("{.x}-{.x+1}")) %>%
  glue_collapse(sep = ', ', last = ', and ')

rp <- report %>% 
  split(list(.$location, .$object_type))

current_analysis$tbl %<>% 
  mutate(across(where(is.numeric), as.integer))

abbrev_key <- c(
  `ACC/AHA` = 'American College of Cardiology and the American Heart Association',
  CI = 'confidence interval',
  BP = 'blood pressure',
  NHANES = 'National Health and Nutrition Examination Survey',
  SBP = 'systolic BP',
  DBP = 'diastolic BP',
  CKD = 'chronic kidney disease',
  CVD = 'cardiovascular disease',
  ASCVD = 'atherosclerotic CVD',
  ACR = 'albumin-to-creatinine ratio',
  HbA1c = 'glycated hemoglobin',
  eGFR = 'estimated glomerular filtration rate'
) %>% 
  enframe(name = 'acronym', value = 'phrase') %>% 
  mutate(counter = 0)

# create global functions that will refer to abbreviations
# in abbrev_key. The first use of the function spells out
# the acronym in full. All uses thereafter just use the
# acronym.

# this object must be defined in order for the abbrev functions 
# to dynamically decide whether to use certain acronyms or not.
# e.g., some journals only want 5 acronyms to be used.

max_abbrevs <- 9

abbrev_funs <- map(
  .x = set_names(abbrev_key$acronym, abbrev_key$acronym),
  .f = ~ {
    function(left_mark = '(', right_mark = ')'){

      if(!exists('max_abbrevs')) max_abbrevs <- nrow(abbrev_key)

      abbrev_index <- which(abbrev_key$acronym == .x)

      low_priority <- abbrev_index > max_abbrevs

      if(low_priority) return(abbrev_key$phrase[abbrev_index])

      counter <- abbrev_key$counter[abbrev_index]

      if(counter < 1){
        # modifies abbrev_key in global environment
        abbrev_key <<- add_counter(abbrev_key, abbrev_index)
        string <- glue_data(abbrev_key[abbrev_index,], 
                            "{phrase} {left_mark}{acronym}{right_mark}")

      } else {

        abbrev_key <<- add_counter(abbrev_key, abbrev_index)
        string <- abbrev_key$acronym[abbrev_index]

      }

      string

    }
  }) %>% 
  clean_list_names(case = 'screaming_snake') %>% 
  list2env(envir = globalenv())

```

Byron C. Jaeger^1^, Swati Sakhuja^2^, Shakia T. Hardy^2^, Philip Akinyelure^2^, Josh Bundy^3^, Paul Muntner^2^, and Paul K. Whelton^3^

1. Department of Biostatistics, University of Alabama at Birmingham
2. Department of Epidemiology, University of Alabama at Birmingham
3. Department of Epidemiology, Tulane University

\newpage

In November 2017, the `r ACC_AHA()` published a guideline for the prevention, detection, evaluation, and management of high `r BP()` in adults.[@whelton20182017] This guideline recommends using both predicted risk for `r CVD()` and `r BP()` levels to guide the initiation of antihypertensive medication.  All adults with `r SBP()` ≥ 140 mm Hg or `r DBP()` ≥ 90 mm Hg, stage 2 hypertension in the guideline, are recommended to initiate antihypertensive medication. Additionally, adults with `r SBP()` between 130 and 139 mm Hg and/or `r DBP()` between 80 and 89 mm Hg, stage 1 hypertension in the guideline, who have prevalent diabetes mellitus, prevalent `r CKD()`, age ≥ 65 years, or high `r ASCVD()` risk are recommended to initiate antihypertensive medication. High `r ASCVD()` risk is defined as prevalent `r CVD()` or a 10-year predicted risk for `r ASCVD()` ≥ 10%

<br>

The 2017 `r ACC_AHA()` guideline recommends computing 10-year predicted risk for `r ASCVD()` in all adults with hypertension to establish a `r BP()` threshold for treatment.[@whelton20182017] However, it has been suggested that the vast majority of adults with diabetes, CKD, or ≥ 65 years of age are likely to have a 10-year predicted risk for `r ASCVD()` ≥ 10%. The purpose of the current analysis was to estimate the proportion of US adults with diabetes, `r CKD()`, or ≥ 65 years of age that have high `r ASCVD()` risk (i.e., a 10-year predicted risk for `r ASCVD()` ≥ 10% or prevalent `r CVD()`).  A secondary objective was to estimate the proportion of adults with stage 1 hypertension and diabetes, `r CKD()`, or age ≥ 65 years who have high `r ASCVD()` risk, as the 2017 `r ACC_AHA()` `r BP()` guideline recommends initiation of antihypertensive medication for these individuals. To accomplish these goals, we analyzed data from `r english(length(exams))` cycles of the US `r NHANES()`.

<br>

# METHODS

`r NHANES()` was designed to assess the health and nutritional status of the non-institutionalized US population and was conducted by the National Center for Health Statistics of the Centers for Disease Control and Prevention.[@nhanes_home] Since 1999-2000, `r NHANES()` has been conducted in two-year cycles using a multistage probability sampling design to select participants. Each cycle is independent with different participants recruited. For the current analysis, the `r english(length(exams))` cycles conducted in `r nhanes_years` were pooled for analysis.[@nhanes_tutorial_weights] The protocols for each `r NHANES()` cycle were approved by the National Center for Health Statistics of the Centers for Disease Control and Prevention Institutional Review Board. Written informed consent was obtained from each participant. The University of Alabama at Birmingham Institutional Review Board considered the analysis of `r NHANES()` data to be exempt research.

<br>

```{r}

# some inline results for the exclusion paragraph below

n_age_40_to_79 <- current_analysis$tbl %>% 
  filter(label == 'Aged 40-79 years') %>% 
  pull(sample_size) %>% 
  table_value()

n_bp_too_low <- current_analysis$tbl %>% 
  filter(str_detect(label, '^3 or more systolic and diastolic')) %>% 
  pull(n_removed) %>% 
  table_value()

n_missing_pcr <- current_analysis$tbl %>% 
  filter(str_detect(label, '^Complete data for variables')) %>% 
  pull(n_removed) %>% 
  table_value()

n_final <- table_value(min(current_analysis$tbl$sample_size))

```

The current analysis was restricted to adults aged 40 to 79 years of age who completed the `r NHANES()` interview and examination (n = `r n_age_40_to_79`). Participants < 40 or > 79 years of age were not included because the Pooled Cohort risk equations are not recommended to be used in these age ranges.[@goff20142013] Participants who did not have three `r SBP()` and `r DBP()` measurements (n = `r n_bp_too_low`) and those who were missing information on age, race, sex, total and high-density lipoprotein cholesterol, smoking status, diabetes, or `r CKD()` status (n = `r n_missing_pcr`) were excluded. After these exclusions, over the `r english(length(exams))` `r NHANES()` cycles, a total of `r n_final` participants were included in the analysis (`r refer_to('fig_include_exclude')`).

<br>

## Data collection

Data were collected during an in-home interview and a study visit completed at a mobile examination center. Standardized questionnaires were used to assess participants’ age, sex, race/ethnicity, smoking habits, medical history and use of antihypertensive medication, oral glucose lowering medication and insulin. Medical history included questions about whether the participant had been told by a doctor or other health professional that they have had a heart attack, coronary heart disease stroke, or heart failure. Blood and urine samples were collected during the medical examination. Of relevance to the current analysis, serum creatinine, serum glucose and `r HB_A1C()` were measured. Diabetes was defined by fasting serum glucose ≥ `r gluc_cutpoint_fasted` mg/dL, non-fasting glucose ≥ `r gluc_cutpoint_fed` mg/dL, `r HB_A1C()` ≥ `r hba1c_cutpoint`%, or self-reported use of insulin or oral glucose lowering medication. `r capitalize(E_GFR())` was calculated using the Chronic Kidney Disease Epidemiology Collaboration equation.[@levey2009new] Urinary albumin and creatinine levels were measured and used to calculate the `r ACR()`. `r CKD()` was defined by an `r E_GFR()` < `r egfr_cutpoint` ml/min/1.73m^2^ or an `r ACR()` ≥ `r acr_cutpoint` mg/dL. Predicted 10-year risk for `r ASCVD()` was calculated using the Pooled Cohort risk equations for participants without prevalent `r CVD()`.[@goff20142013] High `r ASCVD()` risk was defined as prevalent `r CVD()` or a 10-year predicted `r ASCVD()` risk ≥ 10%.

<br>

## Blood pressure measurements

The same protocol was followed to measure `r SBP()` and `r DBP()` in each `r NHANES()` cycle. After survey participants had rested 5 minutes, their `r BP()` was measured by a trained physician using a mercury sphygmomanometer and an appropriately sized cuff. Three `r BP()` measurements were obtained at 30 second intervals. The mean of all available measurements was used to define `r SBP()` and `r DBP()`. Quality control included re-certification of physicians every quarter with retraining if needed. All physicians participated in annual retraining.

<br>

## Blood pressure categories

<!-- These are hard coded to discourage messing with guideline recs -->

Participants not taking antihypertensive medication were grouped into four non-overlapping categories based on the 2017 `r ACC_AHA()` `r BP()` guideline: Normal `r BP()` (`r SBP()` < 120 mm Hg and `r DBP()` < 80 mm Hg), elevated `r BP()` (`r SBP()` between 120 and 129 mm Hg and `r DBP()` < 80 mm Hg), stage 1 hypertension (`r SBP()` between 130 and 139 mm Hg and/or `r DBP()` between 80 and 89 mm Hg with `r SBP()` < 140 mm Hg and `r DBP()` < 90 mm Hg), stage 2 hypertension (`r SBP()` ≥ 140 mm Hg or `r DBP()` ≥ 90 mm Hg). Participants taking antihypertensive medication were placed in a fifth category.

<br>

## Statistical analysis

Analyses were conducted for the overall population and among participants with diabetes, CKD, ≥ 65 years of age, and for those with at least one of these three conditions. Participant characteristics were summarized as mean with their standard error and percentage for continuous and categorical variables, respectively. The percentage of US adults in each of the five `r BP()` categories (defined in the preceding paragraph) was computed. The 25^th^, 50^th^, and 75^th^ percentile of 10-year predicted risk for `r ASCVD()` and the proportion of participants with high `r ASCVD()` risk were estimated overall and within each `r BP()` category. To assess the extent to which participants with a 10-year predicted risk for `r ASCVD()` < 10% were close to the 10% threshold, the proportion of participants in this subgroup with 10-year predicted risk for `r ASCVD()` from < 2.5%, 2.5% to < 5.0%, 5.0% to < 7.5%, and 7.5% to < 10% was estimated. The probability of having high `r ASCVD()` risk was estimated for each year of age from 40 to 79 years using logistic regression. Analyses of participant characteristics, distribution of 10-year predicted risk for `r ASCVD()`, and probability of having high `r ASCVD()` risk were replicated among participants with stage 1 hypertension.

<br>

```{r}

# R version string with date spelled out in words
R_string <- glue::glue("R version {version$major}.{version$minor}")

date_string <- with(version, paste(year, month, day, sep = '/')) %>% 
  as.POSIXct() %>% 
  format("%B %d, %Y")

R_string_fancy <- glue::glue("{R_string} (released {date_string})")

```

`r NHANES()` sampling weights, which were calculated as the inverse probability of being selected for the survey, were used in all calculations to obtain nationally representative estimates of the non-institutionalized US population. The survey design of `r NHANES()` was also taken into account in all calculations. P-values were two-sided. Data analysis was conducted using `r R_string_fancy` along with a collection of open-source software packages.[@cite_R; @cite_drake; @cite_tidyverse; @cite_table.glue] The first author's GitHub repository (< _Link not yet active so we won't be scooped_ >) provides code to reproduce the current manuscript.

<br>

# RESULTS

Among US adults aged 40 to 79 years in `r min(exams)`-`r max(exams)+1`, the estimated prevalence (95% `r CI(left_mark = '[', right_mark = ']')`) of diabetes and `r CKD()` was `r inline$prevDiabetesOverall` and `r inline$prevCkdOverall`, respectively, `r inline$prevAge65Overall` were estimated to have ≥ 65 years of age, and `r inline$prevAnyOverall` had at least one of these conditions (`r refer_to('tab_characteristics')`). An estimated `r inline$prevS1hOverall` of US adults 40 to 79 years of age had stage 1 hypertension (`r refer_to('tab_bpdist')`). Among those with diabetes, CKD, age ≥ 65 years, and any of these three conditions, the estimated prevalence (95% `r CI()`) of stage 1 hypertension was `r inline$prevS1hDiabetes`, `r inline$prevS1hCkd`, `r inline$prevS1hAge65`, `r inline$prevS1hAny`, respectively. Characteristics of US adults 40 to 79 years of age with stage 1 hypertension, overall and for subgroups defined by diabetes, `r CKD()` and age ≥ 65 years, are presented in Table S1.

<br>

## Predicted 10-year risk for atherosclerotic cardiovascular disease

Among US adults aged 40 to 79 years without prevalent `r CVD()`, the estimated median (25^th^, 75^th^ percentiles) 10-year predicted risk for `r ASCVD()` was `r inline$medianPcrOverall` overall and `r inline$medianPcrDiabetes`, `r inline$medianPcrCkd`, `r inline$medianPcrAge65`, `r inline$medianPcrAny` among those with diabetes, CKD, age ≥ 65 years, and any of these three conditions, respectively (`r refer_to('tab_risk_overall')`; top panel). Among those with stage 1 hypertension, the estimated median (25^th^, 75^th^ percentiles) 10-year predicted risk for `r ASCVD()` was `r inline$medianPcrS1hOverall` overall and `r inline$medianPcrS1hDiabetes`, `r inline$medianPcrS1hCkd`, `r inline$medianPcrS1hAge65`, `r inline$medianPcrS1hAny` among those with diabetes, CKD, age ≥ 65 years, and any of these three conditions, respectively.

<br>

Among US adults aged 40 to 79 years, the estimated percentage (95% `r CI()`) at high ASCVD risk was `r inline$prevHighRiskOverall` overall and `r inline$prevHighRiskDiabetes`, `r inline$prevHighRiskCkd`, `r inline$prevHighRiskAge65`, and `r inline$prevHighRiskAny` for those with diabetes, `r CKD()`, age ≥ 65 years, or any of these three conditions, respectively (`r refer_to('tab_risk_overall')`; bottom panel). Among those with stage 1 hypertension, an estimated `r inline$prevHighRiskS1hOverall` had high `r ASCVD()` risk overall and `r inline$prevHighRiskS1hDiabetes`, `r inline$prevHighRiskS1hCkd`, `r inline$prevHighRiskS1hAge65`, and `r inline$prevHighRiskS1hAny` of those with diabetes, `r CKD()`, age ≥ 65 years, or any of these three conditions had high `r ASCVD()` risk, respectively.

<br>

Among US adults aged 40 to 79 years who were not at high risk for `r ASCVD()`, an estimated `r inline$propLowRiskOverall` had 10-year predicted `r ASCVD()` risk < 5% (`r refer_to('fig_hist_ovrl')`). Among those not at high risk for `r ASCVD()` with diabetes, `r CKD()` age ≥ 65 years, and any of these three conditions, an estimated `r inline$propLowRiskDiabetes`, `r inline$propLowRiskCkd`, `r inline$propLowRiskAge65`, and `r inline$propLowRiskAny` had a 10-year predicted `r ASCVD()` risk < 5%, respectively. Among those not at high risk for `r ASCVD()` and with stage 1 hypertension, an estimated `r inline$propLowRiskS1hOverall` overall and  `r inline$propLowRiskS1hDiabetes`, `r inline$propLowRiskS1hCkd`, and `r inline$propLowRiskS1hAge65` of those with diabetes, CKD, and ≥ 65 years of age had a 10-year predicted `r ASCVD()` risk < 5%, respectively (`r refer_to('fig_hist_stg1')`).

<br>

## Age-specific analysis of high `r ASCVD()` risk

The estimated probability of having high `r ASCVD()` risk increased with older age and exceeded 50% at `r inline$ageHighRiskOvrlNoComorb` years for US adults without diabetes or `r CKD()`, compared with `r inline$ageHighRiskOvrlDiabetes` years for US adults with diabetes and `r inline$ageHighRiskOvrlCkd` years for US adults with CKD (`r refer_to('fig_risk_ovrl')`). Among US adults with stage 1 hypertension, the age at which the estimated probability of having high `r ASCVD()` risk exceeded 50% was `r inline$ageHighRiskS1hNoComorb`, `r inline$ageHighRiskS1hDiabetes`, and `r inline$ageHighRiskS1hCkd` years for US adults without diabetes or CKD, with diabetes, and with CKD, respectively (`r refer_to('fig_risk_stg1')`). The minimum age where the probability of high `r ASCVD()` risk exceeded 50% was not determined for adults ≥ 65 years of age as the probability exceeded 50% at all ages above 65 years.

<br>

# DISCUSSION

In the current study, a majority of US adults aged 40 to 79 years with diabetes, `r CKD()`, age ≥ 65 years, and any of these three conditions had a high risk for `r ASCVD()`, defined by a 10-year predicted `r ASCVD()` risk ≥ 10% or prevalent `r CVD()`. Among those with stage 1 hypertension and ≥ 65 years of age, a substantial proportion were at high `r ASCVD()` risk. However, a substantial proportion of US adults with stage 1 hypertension and diabetes or CKD did not have high `r ASCVD()` risk. Among US adults with diabetes or CKD who were not at high `r ASCVD()` risk, approximately half had a 10-year predicted `r ASCVD()` risk < 5%. In contrast, over 80% of US adults aged ≥ 65 years not at high `r ASCVD()` risk had 10-year predicted risk for `r ASCVD()` between 5% and 10%. The probability of having high `r ASCVD()` risk was age-dependent with over 50% of those with diabetes, `r CKD()`, and with neither of these conditions expected to have high `r ASCVD()` risk at ≥ `r inline$ageHighRiskOvrlDiabetes`, ≥ `r inline$ageHighRiskOvrlCkd`, and ≥ `r inline$ageHighRiskOvrlNoComorb` years of age, respectively.

<br>

The current study estimates that roughly 33% of US adults aged 40 to 79 years take antihypertensive medication, including half of those with diabetes, CKD, or ≥ 65 years of age. Among those with diabetes, `r CKD()`, or age ≥ 65 years, only about 10% had stage 1 hypertension. The 2017 `r ACC_AHA()` `r BP()` guideline recommends all of these adults initiate antihypertensive medication and also recommends computing 10-year predicted `r ASCVD()` risk to establish the `r BP()` threshold for treatment.[@whelton20182017] The heterogeneous distribution of 10-year predicted `r ASCVD()` among those with stage 1 hypertension and diabetes, `r CKD()`, or age ≥ 65 years emphasizes the importance of these two recommendations from the guideline. Prior studies have shown the 2017 `r ACC_AHA()` `r BP()` guideline directs antihypertensive medication to those most likely to receive the largest risk reduction benefit,[@colantonio20182017; @jaeger2019cardiovascular] and computing 10-year predicted risk for `r ASCVD()` can both inform patients and facilitate patient provider discussions on a `r BP()` threshold for treatment.

<br>

Previous studies have shown that the use of predicted `r ASCVD()` risk in addition to `r BP()`, as recommended by the 2017 `r ACC_AHA()` `r BP()` guideline, can direct antihypertensive medication to adults likely to receive the largest risk reduction benefit with treatment.[@colantonio20182017] In an analysis of the Reasons for Geographic and Racial Differences in Stroke study, those with stage 1 hypertension recommended versus not recommended to initiate antihypertensive medication by the 2017 `r ACC_AHA()` `r BP()` guideline were roughly 6 times more likely to experience a `r CVD()` event over 8 years of follow-up. Although a substantial proportion of US adults with stage 1 hypertension and diabetes, `r CKD()`, or ≥ 65 years of age did not have a high `r ASCVD()` risk in the current study, they were more likely to have high `r ASCVD()` risk compared to the overall US population with stage 1 hypertension, and may therefore still benefit from initiating antihypertensive medication.

<br>

Previous randomized trials and meta-analyses have investigated whether lower `r BP()` treatment goals reduce incident `r CVD()` risk in patients with diabetes, `r CKD()`, or ≥ 65 years of age. the Systolic Blood Pressure Intervention Trial, which compared the `r SBP()` treatment target of < 120 mm Hg to < 140 mm Hg among older (mean age of 68 years) adults without diabetes, found lower rates of incident fatal and nonfatal major `r CVD()` events among those with the lower `r SBP()` target.[@sprint2015randomized] A randomized trial of patients with diabetes and baseline `r SBP()` < 140 mm hg and `r DBP()` < 80 mm Hg found lower risk of stroke among participants with more intense `r BP()` treatment goals.[@schrier2002effects] A systematic review and meta-analysis of 123 studies with 613,815 participants found strong support for lowering `r SBP()` to less than 130 mm Hg and providing antihypertensive medication to adults with diabetes, CKD, and various other comorbidities.[@ettehad2016blood] This suggests adults with stage 1 hypertension and diabetes, `r CKD()` or ≥ 65 years of age may obtain greater risk reduction versus the overall population by initiating antihypertensive medication.

<br>

It has been suggested that the vast majority of adults with diabetes, CKD, or ≥ 65 years of age have a 10-year predicted risk for `r ASCVD()` ≥ 10%.[@whelton20182017] Previous research has shown that diabetes, `r CKD()`, and advanced age are each associated with an increased risk for `r ASCVD()` events.[@coresh2004evidence; @chronic2010association; @grundy1999diabetes; @lakatta2002age]. Although the current study suggests that a high proportion of US adults with stage 1 hypertension and diabetes or `r CKD()` do not have 10-year predicted risk for `r ASCVD()` ≥ 10%, others have previously shown that adults with diabetes or `r CKD()` have a high lifetime `r CVD()` risk.[@lloyd2006prediction; @hippisley2010derivation] Age-adjusted estimates for the probability of high `r ASCVD()` risk from the current study echo these findings and suggests that adults with diabetes or `r CKD()` develop high `r ASCVD()` risk at a younger age than adults without these conditions. Prior studies have also found that cumulative exposure to high `r BP()` is associated with increased `r CVD()` risk.[@allen2014blood] Therefore, for younger adults with diabetes or `r CKD()`, early initiation of antihypertensive medication may be an important step towards lowering lifetime `r CVD()` risk. Predicted 10-year risk for `r ASCVD()` should be computed for all adults with hypertension, as the 2017 `r ACC_AHA()` `r BP()` guideline recommends, and lifetime `r CVD()` risk should also be taken into account for younger adults with diabetes or `r CKD()`. Taken together, these risk assessment techniques can inform patients and help guide the decision to initiate antihypertensive medication.

<br>

The current study has a number of strengths. The design of `r NHANES()` allows its results to be weighted to provide results that are representative of the US population. Additionally, `r NHANES()` data are collected following a rigorous protocol by trained study staff.  `r BP()` was measured three times following a standardized protocol. However, the results of this study should be considered in the context of known and potential limitations. `r NHANES()` participants completed only one visit and guidelines recommend using the mean `r BP()` averaged over 2 or more visits. Additionally, since only one measurement of serum creatinine and urine albuminuria were available, `r CKD()` status may have been mis-classified in some adults. A total of `r table_value(sum(current_analysis$data$bp_cat == 'Stage 1 hypertension'))` participants had stage 1 hypertension, and some subgroups of this population based on diabetes, `r CKD()`, and ≥ 65 years of age were small.

<br>

In conclusion, a majority of US adults aged 40 to 79 years with diabetes, `r CKD()` or ≥ 65 years of age had 10-year predicted risk for `r ASCVD()` ≥ 10%. However, a lower percentage of those with stage 1 hypertension and diabetes, `r CKD()` or ≥ 65 years of age had 10-year predicted risk for `r ASCVD()` ≥ 10%. While most older US adults with diabetes or `r CKD()` had high `r ASCVD()` risk, a substantial proportion of younger adults with diabetes or `r CKD()` did not. Results from the current study support computing 10-year predicted `r ASCVD()` risk for all adults with hypertension, as recommended by the 2017 `r ACC_AHA()` `r BP()` guideline.

<!-- MAIN TABLES -->

\newpage

`r rp$main.table$caption[1]` 
`r fp_par(line_spacing = 1)`
`r rp$main.table$object[[1]]`

\newpage

`r rp$main.table$caption[2]` 
`r fp_par(line_spacing = 1)`
`r rp$main.table$object[[2]]`

\newpage

`r rp$main.table$caption[3]` 
`r fp_par(line_spacing = 1)`
`r rp$main.table$object[[3]]`

\newpage

<!-- MAIN FIGURES -->

`r rp$main.figure$caption[1]`
`r fp_par(line_spacing = 1)`

```{r fig.width=6, fig.height = 7, dpi = 300}

rp$main.figure$object[[1]]

```

`r rp$main.figure$legend[1]`
`r fp_par(line_spacing = 1)`

\newpage

`r rp$main.figure$caption[2]`
`r fp_par(line_spacing = 1)`

```{r fig.width=6, fig.height = 7, dpi=300}

rp$main.figure$object[[2]]

```

`r rp$main.figure$legend[2]`
`r fp_par(line_spacing = 1)`

\newpage

__SUPPLEMENT__

<!-- SUPPLEMENTAL TABLES -->

\newpage

`r rp$supplement.table$caption[1]` 
`r fp_par(line_spacing = 1)`
`r rp$supplement.table$object[[1]]`

<!-- SUPPLEMENTAL FIGURES -->

\newpage

`r rp$supplement.figure$caption[1]`
`r fp_par(line_spacing = 1)`

```{r out.width='100%'}

knitr::include_graphics('../fig/include_exclude.png')

```

`r rp$supplement.figure$legend[1]`
BP = blood pressure; NHANES = National Health and Nutrition Examination Survey. 
`r fp_par(line_spacing = 1)`

\newpage

`r rp$supplement.figure$caption[2]`
`r fp_par(line_spacing = 1)`

```{r fig.width=6, fig.height = 7, dpi=300}

rp$supplement.figure$object[[2]]

```

`r rp$supplement.figure$legend[2]`
`r fp_par(line_spacing = 1)`

\newpage

`r rp$supplement.figure$caption[3]`
`r fp_par(line_spacing = 1)`

```{r fig.width=6, fig.height = 7, dpi=300}

rp$supplement.figure$object[[3]]

```

`r rp$supplement.figure$legend[3]`
`r fp_par(line_spacing = 1)`

\newpage

# REFERENCES